---
title: "Ploidy change patients scoring"
author: "Philip Smith"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

## Initial steps

### knitr options

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load libraries

```{r load_libraries}
# Load libraries
suppressPackageStartupMessages(suppressWarnings(library(tidyverse)))
suppressPackageStartupMessages(suppressWarnings(library(kableExtra)))
suppressPackageStartupMessages(suppressWarnings(library(Biobase)))
suppressPackageStartupMessages(suppressWarnings(library(QDNAseqmod)))
```

### Set functions

```{r funcs}
depthtocn<-function(x,purity,seqdepth) #converts readdepth to copy number given purity and single copy depth
{
    (x/seqdepth-2*(1-purity))/purity
}

#Plot absolute CN fits for assessment
plotProfile <- function(data = NULL,sample = NULL,ploidy = NULL,purity = NULL){
  x <- data[, sample]
  
  cn <- assayDataElement(x,"copynumber")
  seg <- assayDataElement(x,"segmented")
  
  rel_ploidy <- mean(cn,na.rm=T)
  cellploidy <- ploidy*purity+2*(1-purity)
  seqdepth <- rel_ploidy/cellploidy
  
  temp <- x
  
  abs_cn <- depthtocn(cn,purity,seqdepth)
  abs_seg <- depthtocn(seg,purity,seqdepth)
  
  assayDataElement(temp,"copynumber") <- abs_cn
  assayDataElement(temp,"segmented") <- abs_seg
  
  # plot   
  if(ploidy>5){
    yrange=15
  } else {
    yrange=10
  }
  plot(temp,doCalls=FALSE,showSD=TRUE,logTransform=FALSE,ylim=c(0,yrange),ylab="Absolute tumour CN",
         main=paste(sample," pu=",round(purity,2),
                    " pl=",round(ploidy,2),
                    sep=""),cex.main=0.8)
  abline(h=1:9,col = "blue")
}

```

### load data

```{r metadata}
meta.fitting <- read.table("../../absolute_PRE_down_sampling/fit_QC_predownsample_withsmooth.tsv",header = T,sep = "\t")
ploidy_change_patients <- read.table(file = "data/ploidy_change_pats.tsv",header = T,sep = "\t")
ploidy_change_patients <- as.character(unique(ploidy_change_patients$PATIENT_ID))
```

```{r cn_data}
cnData <- readRDS("../../absolute_PRE_down_sampling/britroc_smoothed_copyNumbersSegmented.rds")
```

```{r cat_vectors}
pats <- c()
cat1 <- c()
cat2 <- c()
cat3 <- c()
notes <- c()
star_rating <- c()
```

### Rating criteria

- Category 1: Selected fits have the highest scoring quantitative quality metrics compared to other sufficiently powered copy number fits (clonality error, TP53 estimate).
- Category 2: No underpowered fits with otherwise acceptable or better quality metrics are available which would contradict the selected copy number fit.
- Category 3: Additional samples, attributed to either diagnosis or relapse groups, support the ploidy change by also conforming to criteria 1 and/or criteria 2.

#### Patient 1 - BRITROC-209

```{r pat1_cat1}
pat <- ploidy_change_patients[1]
pats <- append(pats,pat)
meta.fitting[meta.fitting$PATIENT_ID == pat,-c(2,6,11:15)]
meta.fitting[meta.fitting$PATIENT_ID == pat & meta.fitting$use == "TRUE",-c(2,6,11:15)]
notes <- append(notes,unique(paste0(meta.fitting$notes[meta.fitting$PATIENT_ID == pat],collapse = "")))

cat1 <- append(cat1,"TRUE")
```

```{r pat1_plots}
for(i in 1:nrow(meta.fitting[meta.fitting$PATIENT_ID == pat,-c(2,6,11:15)])){
  plotProfile(data = cnData,
              sample = as.character(meta.fitting$SAMPLE_ID[meta.fitting$PATIENT_ID == pat][i]),
              ploidy = as.numeric(meta.fitting$ploidy[meta.fitting$PATIENT_ID == pat][i]),
              purity = as.numeric(meta.fitting$purity[meta.fitting$PATIENT_ID == pat][i]))
  
}

plotProfile(data = cnData,sample = "IM_295",ploidy = 1.7,purity = 0.56) # Selected
plotProfile(data = cnData,sample = "JBLAB-4960",ploidy = 3.6,purity = 0.82) # Selected
```

```{r pat1_cat2}
IM_295_clonality <- read.table("../../absolute_PRE_down_sampling/clonality_results/IM_295_clonality.csv",sep = ",",header = T) %>%
                      mutate(TP53freq = unique(meta.fitting$TP53freq[meta.fitting$SAMPLE_ID == "IM_295"])) %>%
                      filter(powered == 0) %>%
                      filter(TP53cn > 0) %>%
                      mutate(TPdiff = abs(TP53freq - expected_TP53_AF))

IM_295_clonality[IM_295_clonality$ploidy > 3.2 & IM_295_clonality$ploidy < 4,] %>%
  arrange(TPdiff,clonality)
plotProfile(data = cnData,sample = "IM_295",ploidy = 3.8,purity = 0.31)

IM_295_clonality[IM_295_clonality$ploidy > 5 & IM_295_clonality$ploidy < 6,] %>%
  arrange(TPdiff,clonality)

plotProfile(data = cnData,sample = "IM_295",ploidy = 5.3,purity = 0.28)
paste0("Potential underpowered fit")

JBLAB_4960_clonality <- read.table("../../absolute_PRE_down_sampling/clonality_results/JBLAB-4960_clonality.csv",sep = ",",header = T) %>%
                      mutate(TP53freq = unique(meta.fitting$TP53freq[meta.fitting$SAMPLE_ID == "JBLAB-4960"])) %>%
                      filter(powered == 0) %>%
                      filter(TP53cn > 0) %>%
                      mutate(TPdiff = abs(TP53freq - expected_TP53_AF))

JBLAB_4960_clonality[JBLAB_4960_clonality$ploidy > 1.5 & JBLAB_4960_clonality$ploidy < 2,] %>%
  arrange(TPdiff,clonality)
plotProfile(data = cnData,sample = "JBLAB-4960",ploidy = 1.9,purity = 0.44)

cat2 <- append(cat2,"FALSE")
cat3 <- append(cat3,"TRUE")
star_rating <- append(star_rating,c("**"))
```

#### Patient 2 - BRITROC-216

```{r pat2_cat1}
pat <- ploidy_change_patients[2]
pats <- append(pats,pat)
meta.fitting[meta.fitting$PATIENT_ID == pat,-c(2,6,11:15)]
meta.fitting[meta.fitting$PATIENT_ID == pat & meta.fitting$use == "TRUE",-c(2,6,11:15)]
notes <- append(notes,unique(paste0(meta.fitting$notes[meta.fitting$PATIENT_ID == pat],collapse = "")))

cat1 <- append(cat1,"TRUE")
```

```{r pat2_plots}
for(i in 1:nrow(meta.fitting[meta.fitting$PATIENT_ID == pat,-c(2,6,11:15)])){
  plotProfile(data = cnData,
              sample = as.character(meta.fitting$SAMPLE_ID[meta.fitting$PATIENT_ID == pat][i]),
              ploidy = as.numeric(meta.fitting$ploidy[meta.fitting$PATIENT_ID == pat][i]),
              purity = as.numeric(meta.fitting$purity[meta.fitting$PATIENT_ID == pat][i]))
  
}

plotProfile(data = cnData,sample = "IM_340",ploidy = 1.9,purity = 0.83) # Selected
plotProfile(data = cnData,sample = "IM_341",ploidy = 1.8,purity = 0.84) # Selected
plotProfile(data = cnData,sample = "JBLAB-4965",ploidy = 3.5,purity = 0.78) # Selected
```

```{r pat2_cat2}
JBLAB_4965_clonality <- read.table("../../absolute_PRE_down_sampling/clonality_results/JBLAB-4965_clonality.csv",sep = ",",header = T) %>%
                      mutate(TP53freq = unique(meta.fitting$TP53freq[meta.fitting$SAMPLE_ID == "JBLAB-4965"])) %>%
                      filter(powered == 0) %>%
                      filter(TP53cn > 0) %>%
                      mutate(TPdiff = abs(TP53freq - expected_TP53_AF))

JBLAB_4965_clonality[JBLAB_4965_clonality$ploidy > 1.5 & JBLAB_4965_clonality$ploidy < 2.4,] %>%
  arrange(TPdiff,clonality)
plotProfile(data = cnData,sample = "JBLAB-4965",ploidy = 2.3,purity = 0.44)

IM_340_clonality <- read.table("../../absolute_PRE_down_sampling/clonality_results/IM_340_clonality.csv",sep = ",",header = T) %>%
                      mutate(TP53freq = unique(meta.fitting$TP53freq[meta.fitting$SAMPLE_ID == "IM_340"])) %>%
                      filter(powered == 0) %>%
                      filter(TP53cn > 0) %>%
                      mutate(TPdiff = abs(TP53freq - expected_TP53_AF))

IM_340_clonality[IM_340_clonality$ploidy > 3 & IM_340_clonality$ploidy < 3.8,] %>%
  arrange(TPdiff,clonality)
plotProfile(data = cnData,sample = "IM_340",ploidy = 3.7,purity = 0.66)
paste0("Potential underpowered fit")

IM_341_clonality <- read.table("../../absolute_PRE_down_sampling/clonality_results/IM_341_clonality.csv",sep = ",",header = T) %>%
                      mutate(TP53freq = unique(meta.fitting$TP53freq[meta.fitting$SAMPLE_ID == "IM_341"])) %>%
                      filter(powered == 0) %>%
                      filter(TP53cn > 0) %>%
                      mutate(TPdiff = abs(TP53freq - expected_TP53_AF))

IM_341_clonality[IM_341_clonality$ploidy > 3 & IM_341_clonality$ploidy < 3.8,] %>%
  arrange(TPdiff,clonality)
plotProfile(data = cnData,sample = "IM_341",ploidy = 3.3,purity = 0.84)

cat2 <- append(cat2,"FALSE")
cat3 <- append(cat3,"FALSE")

star_rating <- append(star_rating,c("*"))
```

#### Patient 3 - BRITROC-23

```{r pat3_cat1}
pat <- ploidy_change_patients[3]
pats <- append(pats,pat)
meta.fitting[meta.fitting$PATIENT_ID == pat,-c(2,6,11:15)]
meta.fitting[meta.fitting$PATIENT_ID == pat & meta.fitting$use == "TRUE",-c(2,6,11:15)]
notes <- append(notes,unique(paste0(meta.fitting$notes[meta.fitting$PATIENT_ID == pat],collapse = "")))

cat1 <- append(cat1,"TRUE")
```

```{r pat3_plots}
for(i in 1:nrow(meta.fitting[meta.fitting$PATIENT_ID == pat,-c(2,6,11:15)])){
  plotProfile(data = cnData,
              sample = as.character(meta.fitting$SAMPLE_ID[meta.fitting$PATIENT_ID == pat][i]),
              ploidy = as.numeric(meta.fitting$ploidy[meta.fitting$PATIENT_ID == pat][i]),
              purity = as.numeric(meta.fitting$purity[meta.fitting$PATIENT_ID == pat][i]))
  
}

plotProfile(data = cnData,sample = "IM_56",ploidy = 1.9,purity = 0.56) # Selected
plotProfile(data = cnData,sample = "JBLAB-4128",ploidy = 3.6,purity = 0.82) # Selected
```

```{r pat3_cat2}
IM_56_clonality <- read.table("../../absolute_PRE_down_sampling/clonality_results/IM_56_clonality.csv",sep = ",",header = T) %>%
                      mutate(TP53freq = unique(meta.fitting$TP53freq[meta.fitting$SAMPLE_ID == "IM_56"])) %>%
                      filter(powered == 0) %>%
                      filter(TP53cn > 0) %>%
                      mutate(TPdiff = abs(TP53freq - expected_TP53_AF))

IM_56_clonality[IM_56_clonality$ploidy > 3.2 & IM_56_clonality$ploidy < 4,] %>%
  arrange(TPdiff,clonality)
plotProfile(data = cnData,sample = "IM_56",ploidy = 3.5,purity = 0.44)
paste0("Potential underpowered fit")

JBLAB_4128_clonality <- read.table("../../absolute_PRE_down_sampling/clonality_results/JBLAB-4128_clonality.csv",sep = ",",header = T) %>%
                      mutate(TP53freq = unique(meta.fitting$TP53freq[meta.fitting$SAMPLE_ID == "JBLAB-4128"])) %>%
                      filter(powered == 0) %>%
                      filter(TP53cn > 0) %>%
                      mutate(TPdiff = abs(TP53freq - expected_TP53_AF))

JBLAB_4128_clonality[JBLAB_4128_clonality$ploidy > 1.5 & JBLAB_4128_clonality$ploidy < 2.4,] %>%
  arrange(TPdiff,clonality)
plotProfile(data = cnData,sample = "JBLAB-4128",ploidy = 2.3,purity = 0.59)


cat2 <- append(cat2,"FALSE")
cat3 <- append(cat3,"TRUE")

star_rating <- append(star_rating,c("**"))
```

#### Patient 4 - BRITROC-241

```{r pat4_cat1}
pat <- ploidy_change_patients[4]
pats <- append(pats,pat)
meta.fitting[meta.fitting$PATIENT_ID == pat,-c(2,6,11:15)]
meta.fitting[meta.fitting$PATIENT_ID == pat & meta.fitting$use == "TRUE",-c(2,6,11:15)]
notes <- append(notes,unique(paste0(meta.fitting$notes[meta.fitting$PATIENT_ID == pat],collapse = "")))

cat1 <- append(cat1,"TRUE")
```

```{r pat4_plots}
for(i in 1:nrow(meta.fitting[meta.fitting$PATIENT_ID == pat,-c(2,6,11:15)])){
  plotProfile(data = cnData,
              sample = as.character(meta.fitting$SAMPLE_ID[meta.fitting$PATIENT_ID == pat][i]),
              ploidy = as.numeric(meta.fitting$ploidy[meta.fitting$PATIENT_ID == pat][i]),
              purity = as.numeric(meta.fitting$purity[meta.fitting$PATIENT_ID == pat][i]))
  
}

plotProfile(data = cnData,sample = "IM_423",ploidy = 1.8,purity = 0.70) # Selected
plotProfile(data = cnData,sample = "JBLAB-4996",ploidy = 3.0,purity = 0.76) # Selected
```

```{r pat4_cat2}
JBLAB_4996_clonality <- read.table("../../absolute_PRE_down_sampling/clonality_results/JBLAB-4996_clonality.csv",sep = ",",header = T) %>%
                      mutate(TP53freq = unique(meta.fitting$TP53freq[meta.fitting$SAMPLE_ID == "JBLAB-4996"])) %>%
                      filter(powered == 0) %>%
                      filter(TP53cn > 0) %>%
                      mutate(TPdiff = abs(TP53freq - expected_TP53_AF))

JBLAB_4996_clonality[JBLAB_4996_clonality$ploidy > 1.6 & JBLAB_4996_clonality$ploidy < 2.2,] %>%
  arrange(TPdiff,clonality)
plotProfile(data = cnData,sample = "JBLAB-4996",ploidy = 2.0,purity = 0.48)

IM_423_clonality <- read.table("../../absolute_PRE_down_sampling/clonality_results/IM_423_clonality.csv",sep = ",",header = T) %>%
                      mutate(TP53freq = unique(meta.fitting$TP53freq[meta.fitting$SAMPLE_ID == "IM_423"])) %>%
                      filter(powered == 0) %>%
                      filter(TP53cn > 0) %>%
                      mutate(TPdiff = abs(TP53freq - expected_TP53_AF))

IM_423_clonality[IM_423_clonality$ploidy > 2.7 & IM_423_clonality$ploidy < 3.4,] %>%
  arrange(TPdiff,clonality)
plotProfile(data = cnData,sample = "IM_423",ploidy = 3.3,purity = 0.54)


cat2 <- append(cat2,"TRUE")
cat3 <- append(cat3,"FALSE")

star_rating <- append(star_rating,c("**"))
```

#### Patient 5 - BRITROC-248

```{r pat5_cat1}
pat <- ploidy_change_patients[5]
pats <- append(pats,pat)
meta.fitting[meta.fitting$PATIENT_ID == pat,-c(2,6,11:15)]
meta.fitting[meta.fitting$PATIENT_ID == pat & meta.fitting$use == "TRUE",-c(2,6,11:15)]
notes <- append(notes,unique(paste0(meta.fitting$notes[meta.fitting$PATIENT_ID == pat],collapse = "")))

cat1 <- append(cat1,"TRUE")
```

```{r pat5_plots}
for(i in 1:nrow(meta.fitting[meta.fitting$PATIENT_ID == pat,-c(2,6,11:15)])){
  plotProfile(data = cnData,
              sample = as.character(meta.fitting$SAMPLE_ID[meta.fitting$PATIENT_ID == pat][i]),
              ploidy = as.numeric(meta.fitting$ploidy[meta.fitting$PATIENT_ID == pat][i]),
              purity = as.numeric(meta.fitting$purity[meta.fitting$PATIENT_ID == pat][i]))
  
}

plotProfile(data = cnData,sample = "IM_403",ploidy = 5.3,purity = 0.48) # Selected
plotProfile(data = cnData,sample = "JBLAB-19302",ploidy = 2.7,purity = 0.44) # Selected
plotProfile(data = cnData,sample = "JBLAB-19303",ploidy = 2.6,purity = 0.47) # Selected
```

```{r pat5_cat2}
JBLAB_19302_clonality <- read.table("../../absolute_PRE_down_sampling/clonality_results/JBLAB-19302_clonality.csv",sep = ",",header = T) %>%
                      mutate(TP53freq = unique(meta.fitting$TP53freq[meta.fitting$SAMPLE_ID == "JBLAB-19302"])) %>%
                      filter(powered == 0) %>%
                      filter(TP53cn > 0) %>%
                      mutate(TPdiff = abs(TP53freq - expected_TP53_AF))

JBLAB_19302_clonality[JBLAB_19302_clonality$ploidy > 4.8 & JBLAB_19302_clonality$ploidy < 5.5,] %>%
  arrange(TPdiff,clonality)
plotProfile(data = cnData,sample = "JBLAB-19302",ploidy = 5.3,purity = 0.29)
paste0("Potential underpowered fit")

JBLAB_19303_clonality <- read.table("../../absolute_PRE_down_sampling/clonality_results/JBLAB-19303_clonality.csv",sep = ",",header = T) %>%
                      mutate(TP53freq = unique(meta.fitting$TP53freq[meta.fitting$SAMPLE_ID == "JBLAB-19303"])) %>%
                      filter(powered == 0) %>%
                      filter(TP53cn > 0) %>%
                      mutate(TPdiff = abs(TP53freq - expected_TP53_AF))

JBLAB_19303_clonality[JBLAB_19303_clonality$ploidy > 4.7 & JBLAB_19303_clonality$ploidy < 5.4,] %>%
  arrange(TPdiff,clonality)
plotProfile(data = cnData,sample = "JBLAB-19303",ploidy = 5.0,purity = 0.28)

IM_403_clonality <- read.table("../../absolute_PRE_down_sampling/clonality_results/IM_403_clonality.csv",sep = ",",header = T) %>%
                      mutate(TP53freq = unique(meta.fitting$TP53freq[meta.fitting$SAMPLE_ID == "IM_403"])) %>%
                      filter(powered == 0) %>%
                      filter(TP53cn > 0) %>%
                      mutate(TPdiff = abs(TP53freq - expected_TP53_AF))

IM_403_clonality[IM_403_clonality$ploidy > 2.4 & IM_403_clonality$ploidy < 3.3,] %>%
  arrange(TPdiff,clonality)
plotProfile(data = cnData,sample = "IM_403",ploidy = 3.1,purity = 0.41)

cat2 <- append(cat2,"FALSE")
cat3 <- append(cat3,"TRUE")

star_rating <- append(star_rating,c("**"))
```

#### Patient 6 - BRITROC-267

```{r pat6_cat1}
pat <- ploidy_change_patients[6]
pats <- append(pats,pat)
meta.fitting[meta.fitting$PATIENT_ID == pat,-c(2,6,11:15)]
meta.fitting[meta.fitting$PATIENT_ID == pat & meta.fitting$use == "TRUE",-c(2,6,11:15)]
notes <- append(notes,unique(paste0(meta.fitting$notes[meta.fitting$PATIENT_ID == pat],collapse = "")))

cat1 <- append(cat1,"TRUE")
```

```{r pat6_plots}
for(i in 1:nrow(meta.fitting[meta.fitting$PATIENT_ID == pat,-c(2,6,11:15)])){
  plotProfile(data = cnData,
              sample = as.character(meta.fitting$SAMPLE_ID[meta.fitting$PATIENT_ID == pat][i]),
              ploidy = as.numeric(meta.fitting$ploidy[meta.fitting$PATIENT_ID == pat][i]),
              purity = as.numeric(meta.fitting$purity[meta.fitting$PATIENT_ID == pat][i]))
  
}

plotProfile(data = cnData,sample = "IM_383",ploidy = 1.8,purity = 0.63) # Selected
plotProfile(data = cnData,sample = "JBLAB-19330",ploidy = 3.4,purity = 0.65) # Selected
```

```{r pat6_cat2}
IM_383_clonality <- read.table("../../absolute_PRE_down_sampling/clonality_results/IM_383_clonality.csv",sep = ",",header = T) %>%
                      mutate(TP53freq = unique(meta.fitting$TP53freq[meta.fitting$SAMPLE_ID == "IM_383"])) %>%
                      filter(powered == 0) %>%
                      filter(TP53cn > 0) %>%
                      mutate(TPdiff = abs(TP53freq - expected_TP53_AF))

IM_383_clonality[IM_383_clonality$ploidy > 3.0 & IM_383_clonality$ploidy < 3.8,] %>%
  arrange(TPdiff,clonality)
plotProfile(data = cnData,sample = "IM_383",ploidy = 3.7,purity = 0.41)

JBLAB_19330_clonality <- read.table("../../absolute_PRE_down_sampling/clonality_results/JBLAB-19330_clonality.csv",sep = ",",header = T) %>%
                      mutate(TP53freq = unique(meta.fitting$TP53freq[meta.fitting$SAMPLE_ID == "JBLAB-19330"])) %>%
                      filter(powered == 0) %>%
                      filter(TP53cn > 0) %>%
                      mutate(TPdiff = abs(TP53freq - expected_TP53_AF))

JBLAB_19330_clonality[JBLAB_19330_clonality$ploidy > 1.5 & JBLAB_19330_clonality$ploidy < 2.3,] %>%
  arrange(TPdiff,clonality)
plotProfile(data = cnData,sample = "JBLAB-19330",ploidy = 2.2,purity = 0.58)

cat2 <- append(cat2,"TRUE")
cat3 <- append(cat3,"FALSE")

star_rating <- append(star_rating,c("**"))
```

#### Patient 7 - BRITROC-274

```{r pat7_cat1}
pat <- ploidy_change_patients[7]
pats <- append(pats,pat)
meta.fitting[meta.fitting$PATIENT_ID == pat,-c(2,6,11:15)]
meta.fitting[meta.fitting$PATIENT_ID == pat & meta.fitting$use == "TRUE",-c(2,6,11:15)]
notes <- append(notes,unique(paste0(meta.fitting$notes[meta.fitting$PATIENT_ID == pat],collapse = "")))

cat1 <- append(cat1,"TRUE")
```

```{r pat7_plots}
for(i in 1:nrow(meta.fitting[meta.fitting$PATIENT_ID == pat,-c(2,6,11:15)])){
  if(as.character(meta.fitting$SAMPLE_ID[meta.fitting$PATIENT_ID == pat][i]) == "IM_397"){
    next() # IM_397 is a non-HGSOC sample
  }
  plotProfile(data = cnData,
              sample = as.character(meta.fitting$SAMPLE_ID[meta.fitting$PATIENT_ID == pat][i]),
              ploidy = as.numeric(meta.fitting$ploidy[meta.fitting$PATIENT_ID == pat][i]),
              purity = as.numeric(meta.fitting$purity[meta.fitting$PATIENT_ID == pat][i]))
  
}

plotProfile(data = cnData,sample = "IM_395",ploidy = 1.8,purity = 0.63) # Selected
plotProfile(data = cnData,sample = "IM_396",ploidy = 1.8,purity = 0.70) # Selected
plotProfile(data = cnData,sample = "JBLAB-19337",ploidy = 1.9,purity = 0.67) # Selected
plotProfile(data = cnData,sample = "JBLAB-19338",ploidy = 3.9,purity = 1.00) # Selected
```

```{r pat7_cat2}
JBLAB_19338_clonality <- read.table("../../absolute_PRE_down_sampling/clonality_results/JBLAB-19338_clonality.csv",sep = ",",header = T) %>%
                      mutate(TP53freq = unique(meta.fitting$TP53freq[meta.fitting$SAMPLE_ID == "JBLAB-19338"])) %>%
                      filter(powered == 0) %>%
                      filter(TP53cn > 0) %>%
                      mutate(TPdiff = abs(TP53freq - expected_TP53_AF))

JBLAB_19338_clonality[JBLAB_19338_clonality$ploidy > 1.5 & JBLAB_19338_clonality$ploidy < 2.4,] %>%
  arrange(TPdiff,clonality)
plotProfile(data = cnData,sample = "JBLAB-19338",ploidy = 3.7,purity = 0.41)

JBLAB_19337_clonality <- read.table("../../absolute_PRE_down_sampling/clonality_results/JBLAB-19337_clonality.csv",sep = ",",header = T) %>%
                      mutate(TP53freq = unique(meta.fitting$TP53freq[meta.fitting$SAMPLE_ID == "JBLAB-19337"])) %>%
                      filter(powered == 0) %>%
                      filter(TP53cn > 0) %>%
                      mutate(TPdiff = abs(TP53freq - expected_TP53_AF))

JBLAB_19337_clonality[JBLAB_19337_clonality$ploidy > 3.5 & JBLAB_19337_clonality$ploidy < 4.4,] %>%
  arrange(TPdiff,clonality)
plotProfile(data = cnData,sample = "JBLAB-19337",ploidy = 3.8,purity = 0.48)
paste0("Potential underpowered fit")

IM_395_clonality <- read.table("../../absolute_PRE_down_sampling/clonality_results/IM_395_clonality.csv",sep = ",",header = T) %>%
                      mutate(TP53freq = unique(meta.fitting$TP53freq[meta.fitting$SAMPLE_ID == "IM_395"])) %>%
                      filter(powered == 0) %>%
                      filter(TP53cn > 0) %>%
                      mutate(TPdiff = abs(TP53freq - expected_TP53_AF))

IM_395_clonality[IM_395_clonality$ploidy > 3.5 & IM_395_clonality$ploidy < 4.4,] %>%
  arrange(TPdiff,clonality)
plotProfile(data = cnData,sample = "IM_395",ploidy = 3.6,purity = 0.57)
paste0("Potential underpowered fit")

IM_396_clonality <- read.table("../../absolute_PRE_down_sampling/clonality_results/IM_396_clonality.csv",sep = ",",header = T) %>%
                      mutate(TP53freq = unique(meta.fitting$TP53freq[meta.fitting$SAMPLE_ID == "IM_396"])) %>%
                      filter(powered == 0) %>%
                      filter(TP53cn > 0) %>%
                      mutate(TPdiff = abs(TP53freq - expected_TP53_AF))

IM_396_clonality[IM_396_clonality$ploidy > 3.3 & IM_396_clonality$ploidy < 4.4,] %>%
  arrange(TPdiff,clonality)
plotProfile(data = cnData,sample = "IM_396",ploidy = 4.3,purity = 0.40)

cat2 <- append(cat2,"FALSE")
cat3 <- append(cat3,"FALSE")

star_rating <- append(star_rating,c("*"))
```

#### Patient 8 - BRITROC-67

```{r pat8_cat1}
pat <- ploidy_change_patients[8]
pats <- append(pats,pat)
meta.fitting[meta.fitting$PATIENT_ID == pat,-c(2,6,11:15)]
meta.fitting[meta.fitting$PATIENT_ID == pat & meta.fitting$use == "TRUE",-c(2,6,11:15)]
notes <- append(notes,unique(paste0(meta.fitting$notes[meta.fitting$PATIENT_ID == pat],collapse = "")))

cat1 <- append(cat1,"TRUE")
```

```{r pat8_plots}
for(i in 1:nrow(meta.fitting[meta.fitting$PATIENT_ID == pat,-c(2,6,11:15)])){
  plotProfile(data = cnData,
              sample = as.character(meta.fitting$SAMPLE_ID[meta.fitting$PATIENT_ID == pat][i]),
              ploidy = as.numeric(meta.fitting$ploidy[meta.fitting$PATIENT_ID == pat][i]),
              purity = as.numeric(meta.fitting$purity[meta.fitting$PATIENT_ID == pat][i]))
  
}

plotProfile(data = cnData,sample = "IM_115",ploidy = 2.9,purity = 1.0) # Selected
plotProfile(data = cnData,sample = "JBLAB-4179",ploidy = 1.9,purity = 0.88) # Selected
```

```{r pat8_cat2}
IM_115_clonality <- read.table("../../absolute_PRE_down_sampling/clonality_results/IM_115_clonality.csv",sep = ",",header = T) %>%
                      mutate(TP53freq = unique(meta.fitting$TP53freq[meta.fitting$SAMPLE_ID == "IM_115"])) %>%
                      filter(powered == 0) %>%
                      filter(TP53cn > 0) %>%
                      mutate(TPdiff = abs(TP53freq - expected_TP53_AF))

IM_115_clonality[IM_115_clonality$ploidy > 1.5 & IM_115_clonality$ploidy < 2.4,] %>%
  arrange(TPdiff,clonality)
plotProfile(data = cnData,sample = "IM_115",ploidy = 3.7,purity = 0.41)

JBLAB_4179_clonality <- read.table("../../absolute_PRE_down_sampling/clonality_results/JBLAB-4179_clonality.csv",sep = ",",header = T) %>%
                      mutate(TP53freq = unique(meta.fitting$TP53freq[meta.fitting$SAMPLE_ID == "JBLAB-4179"])) %>%
                      filter(powered == 0) %>%
                      filter(TP53cn > 0) %>%
                      mutate(TPdiff = abs(TP53freq - expected_TP53_AF))

JBLAB_4179_clonality[JBLAB_4179_clonality$ploidy > 2.6 & JBLAB_4179_clonality$ploidy < 3.4,] %>%
  arrange(TPdiff,clonality)
plotProfile(data = cnData,sample = "JBLAB-4179",ploidy = 3.7,purity = 0.41)

cat2 <- append(cat2,"TRUE")
cat3 <- append(cat3,"FALSE")

star_rating <- append(star_rating,c("**"))
```

#### Patient 9 - BRITROC-74

```{r pat9_cat1}
pat <- ploidy_change_patients[9]
pats <- append(pats,pat)
meta.fitting[meta.fitting$PATIENT_ID == pat,-c(2,6,11:15)]
meta.fitting[meta.fitting$PATIENT_ID == pat & meta.fitting$use == "TRUE",-c(2,6,11:15)]
notes <- append(notes,unique(paste0(meta.fitting$notes[meta.fitting$PATIENT_ID == pat],collapse = "")))

cat1 <- append(cat1,"TRUE")
```

```{r pat9_plots}
for(i in 1:nrow(meta.fitting[meta.fitting$PATIENT_ID == pat,-c(2,6,11:15)])){
  plotProfile(data = cnData,
              sample = as.character(meta.fitting$SAMPLE_ID[meta.fitting$PATIENT_ID == pat][i]),
              ploidy = as.numeric(meta.fitting$ploidy[meta.fitting$PATIENT_ID == pat][i]),
              purity = as.numeric(meta.fitting$purity[meta.fitting$PATIENT_ID == pat][i]))
  
}

plotProfile(data = cnData,sample = "IM_124",ploidy = 2.9,purity = 0.9) # Selected
plotProfile(data = cnData,sample = "JBLAB-4186",ploidy = 3.2,purity = 0.57) # Selected
plotProfile(data = cnData,sample = "JBLAB-4187",ploidy = 3.2,purity = 0.64) # Selected
plotProfile(data = cnData,sample = "JBLAB-4188",ploidy = 3.1,purity = 0.75) # Selected
plotProfile(data = cnData,sample = "JBLAB-4189",ploidy = 3.1,purity = 0.64) # Selected
```

```{r pat9_cat2}
IM_124_clonality <- read.table("../../absolute_PRE_down_sampling/clonality_results/IM_124_clonality.csv",sep = ",",header = T) %>%
                      mutate(TP53freq = unique(meta.fitting$TP53freq[meta.fitting$SAMPLE_ID == "IM_124"])) %>%
                      filter(powered == 0) %>%
                      filter(TP53cn > 0) %>%
                      mutate(TPdiff = abs(TP53freq - expected_TP53_AF))

IM_124_clonality[IM_124_clonality$ploidy > 2.8 & IM_124_clonality$ploidy < 3.8,] %>%
  arrange(TPdiff,clonality)
plotProfile(data = cnData,sample = "IM_115",ploidy = 3.7,purity = 0.62)

cat2 <- append(cat2,"TRUE")
cat3 <- append(cat3,"TRUE")

star_rating <- append(star_rating,c("***"))
```

### Final table

```{r ploidy_table_final}
ploidy_change_scoring_table <- data.frame(patient=pats,
           samples=c(unlist(lapply(pats,FUN = function(x){
            samples <- unique(meta.fitting$SAMPLE_ID[meta.fitting$PATIENT_ID == x])
            colsamples <- paste0(samples,collapse = ",")
            return(colsamples)
            }))),
           `category 1` = cat1,
           `category 2` = cat2,
           `category 3` = cat3,
           `star rating` = star_rating,
           `fit note` = notes,check.names = F)
write.table(ploidy_change_scoring_table,file = "data/ploidy_change_scoring_table.tsv",quote = F,row.names = F,col.names = T,sep = "\t")
ploidy_change_scoring_table
```

```{r sessioninfo}
sessionInfo()
```