---
title: "Generate clinical data"
author: "Philip Smith"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

## Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load libs
```{r libs}
## Load required libraries
library(RPostgres)
library(config)
library(tidyverse)
library(ggnewscale)
library(zoo)
library(survival)
library(survminer)
```

### source colour_palettes
```{r get_colour_palettes}
source("colour_palettes.R")
```

### setup dirs
```{r setup_dir}
if(!dir.exists("plots/")){
  dir.create("plots/")
}

if(!dir.exists("source_data/")){
  dir.create("source_data/")
}
```

### Check for clinDB
```{r clinDb_need}
# Checks if connection to clinicalDB is required
# If generated files already exist then chunks with clinDB = FALSE are skipped
if(!file.exists("britroc_cohort_patient_data.tsv")){
    clinDB_need <- TRUE
} else {
    clinDB_need <- FALSE
}
```

## Load clincal data
```{r db_open,eval=clinDB_need}
## Retrieve clinical database login details
clin_config <- config::get("clinDB")
## Establish connection to Postgres DB
britroc_con <- dbConnect(RPostgres::Postgres(),
                         dbname=clin_config$dbname,
                         host=clin_config$host,
                         port = clin_config$port,
                         user = clin_config$user,
                         password = clin_config$password
)
```

### Clinical data tables
```{r read_tables,eval=clinDB_need}
pat.data <- RPostgres::dbReadTable(britroc_con, 'patients')
chemo_lines <- RPostgres::dbReadTable(britroc_con, 'chemotherapy_lines')
chemo_drugs <- RPostgres::dbReadTable(britroc_con, 'chemotherapy_lines_drugs')
```

### Filter data
```{r filter_tables,eval=clinDB_need}
paste0("Patients with pat.data data: ",length(unique(pat.data$britroc_number)))

chemo_lines <- chemo_lines %>%
  group_by(fk_britroc_number) %>%
  mutate(total_reg_chemo = max(chemotherapy_line)) %>%
  mutate(pre_reg_chemo = max(chemotherapy_line[time_relation_to_image_guided_biopsy == "before"])) %>%
  mutate(post_reg_chemo = total_reg_chemo - pre_reg_chemo)

paste0("Patients with chemo_line data: ",length(unique(chemo_lines$fk_britroc_number)))

## Remove patients with no progression date information (TO BE CHANGED)
chemo_lines <- chemo_lines %>%
  filter(!c(is.na(progression_date) & is.na(progression_date_misc)))

paste0("Patients with chemo_line progression data: ",length(unique(chemo_lines$fk_britroc_number)))

chemo_drugs <- chemo_drugs %>%
                group_by(fk_britroc_number,chemotherapy_line) %>%
                mutate(course = row_number()) %>%
                dplyr::select(-id)

paste0("Patients with chemo_drug data: ",length(unique(chemo_drugs$fk_britroc_number)))
missing_chemo <- sort(unique(c(setdiff(chemo_lines$fk_britroc_number,chemo_drugs$fk_britroc_number),
       setdiff(chemo_drugs$fk_britroc_number,chemo_lines$fk_britroc_number))))
paste0("Patients not in both chemo_tables: ",paste0(missing_chemo,collapse = ","))
```

### Join patient and chemo data
```{r combine_tables,eval=clinDB_need}
chemo_data <- as.data.frame(full_join(x=chemo_lines,y=chemo_drugs,by = c("fk_britroc_number","chemotherapy_line")))
paste0("Patients with chemo data: ",length(unique(chemo_data$fk_britroc_number)))

survival_data <- as.data.frame(full_join(x = pat.data,y = chemo_data, by = c("britroc_number"="fk_britroc_number")) %>%
                    dplyr::select(britroc_number,
                           age,
                           diagnosis_date,
                           diagnosis_date_misc,
                           tumour_stage_at_diagnosis,
                           tumour_substage_at_diagnosis,
                           registration_date,
                           date_of_death,
                           date_alive,
                           progression_date,
                           progression_date_misc,
                           drug,
                           pt_sensitivity_at_reg,
                           total_reg_chemo,
                           pre_reg_chemo,
                           post_reg_chemo) %>%
                    mutate(plat=grepl(x = drug,pattern = "plat",ignore.case = T)) %>%
                    mutate(britroc_number = as.character(britroc_number)) %>%
                    group_by(britroc_number) %>%
                    mutate(drugs = paste0(drug, collapse = ",")) %>%
                    mutate(plat = paste0(plat, collapse = ",")) %>%
                    dplyr::select(-drug) %>%
                    distinct())

paste0("Available patient data: ",length(unique(survival_data$britroc_number)))
```

### Fix missing dates
```{r date_fixes,eval=clinDB_need}
date_cols <- c("diagnosis_date","registration_date","date_of_death","progression_date","date_alive")
survival_data[date_cols] <- lapply(survival_data[date_cols],as.Date, "%m/%d/%Y")

## Split diagnosis_date_misc to save diagnosis dates with month and year data but no day
survival_data$diagnosis_date_misc_month <- str_split(survival_data$diagnosis_date_misc,pattern = " ",n = 2,simplify = T)[,1]
survival_data$diagnosis_date_misc_year <- str_split(survival_data$diagnosis_date_misc,pattern = " ",n = 2,simplify = T)[,2]
## NA converted to characters during database gen - revert to NA characters
survival_data$diagnosis_date_misc_month[which(survival_data$diagnosis_date_misc_month == "NA")] <- NA
## Index usable month/year data
diag_data_idx <- which(is.na(survival_data$diagnosis_date) & !is.na(survival_data$diagnosis_date_misc_month) & !is.na(survival_data$diagnosis_date_misc_year))
## Generate date-format data to add to 
diag_dates <- paste0(survival_data$diagnosis_date_misc_month[diag_data_idx],survival_data$diagnosis_date_misc_year[diag_data_idx])
survival_data$diagnosis_date[diag_data_idx] <- as.Date(as.yearmon(diag_dates, format = "%b%Y"))

## Split progression_date_misc to save progression dates with month and year data but no day
survival_data$progression_date_misc_month <- str_split(survival_data$progression_date_misc,pattern = " ",n = 2,simplify = T)[,1]
survival_data$progression_date_misc_year <- str_split(survival_data$progression_date_misc,pattern = " ",n = 2,simplify = T)[,2]
## NA converted to characters during database gen - revert to NA characters
survival_data$progression_date_misc_month[which(survival_data$progression_date_misc_month == "NA")] <- NA
## Index usable month/year data
prog_data_idx <- which(is.na(survival_data$progression_date) & !is.na(survival_data$progression_date_misc_month) & !is.na(survival_data$progression_date_misc_year))
## Generate date-format data to add to 
prog_dates <- paste0(survival_data$progression_date_misc_month[prog_data_idx],survival_data$progression_date_misc_year[prog_data_idx])
survival_data$progression_date[prog_data_idx] <- as.Date(as.yearmon(prog_dates, format = "%b%Y"))
```

### Fix progression starts
```{r progress_filter,eval=clinDB_need}
missing_prog <- survival_data %>%
                  group_by(britroc_number) %>%
                  mutate(prog_reg_closest = all(is.na(progression_date))) %>%
                  filter(prog_reg_closest)
## Take progression date closest to registration (BEST CURRENT FIX)
survival_data <- survival_data %>%
                  group_by(britroc_number) %>%
                  filter(!is.na(progression_date)) %>%
                  mutate(prog_reg_closest = abs(as.numeric(progression_date - registration_date)) == 
                           min(abs(as.numeric(progression_date - registration_date)),na.rm = T)) %>%
                  filter(prog_reg_closest) %>%
                  dplyr::select(-prog_reg_closest)

paste0("Patients with progression date: ",length(unique(survival_data$britroc_number)))
paste0("Patients missing any progression date: ",paste0(unique(missing_prog$britroc_number),collapse = ", "))
```

### Exclude missing diag dates
```{r exlc_diag,eval=clinDB_need}
missing_diag <- survival_data[is.na(survival_data$diagnosis_date),]
## take date at middle of year
missingDates <- str_split(pattern = " ",
                          survival_data$diagnosis_date_misc[is.na(survival_data$diagnosis_date)],
                          simplify = T,2)[,2]

newDates <- as.Date(paste0(missingDates,"-07-02"))

survival_data$diagnosis_date[is.na(survival_data$diagnosis_date)] <- newDates

paste0("Patients with diagnosis date: ",length(unique(survival_data$britroc_number)))
paste0("Patients missing any diagnosis date: ",paste0(unique(missing_diag$britroc_number),collapse = ", "))
```

### Compute surv metrics
```{r add_surv_metrics,eval=clinDB_need}
## last documented clinical outcome 
survival_data$doc_date <- survival_data$date_alive
survidx <- which(!is.na(survival_data$date_of_death))
survival_data$doc_date[survidx] <- survival_data$date_of_death[survidx]
survival_data$status <- 0
survival_data$status[survidx] <- 1

## Add OS,PFS,time from diag to registration (int_Start), and time from diag to end time point (int_end)
survival_data$os <- difftime(survival_data$doc_date,survival_data$diagnosis_date)
survival_data$pfs <- difftime(survival_data$progression_date,survival_data$diagnosis_date)
survival_data$int_start <- difftime(survival_data$registration_date,survival_data$diagnosis_date)
survival_data$int_end <- difftime(survival_data$doc_date,survival_data$diagnosis_date)
```

### Filter surv metrics
```{r filter_surv_metrics,eval=clinDB_need}
# note odd group of patients with time to progression greater than time to registration - given that progression is the reason for registration
diff_pfs_int <- survival_data[which(survival_data$int_start < survival_data$pfs),] %>%
                  mutate(diff_int_pfs = pfs - int_start)
paste0("Patients with Progression greater than time to registration: ",
       length(unique(survival_data$britroc_number[which(survival_data$int_start < survival_data$pfs)])))
diff_pfs_int[,c(1,23:ncol(diff_pfs_int))]
```

### Write clinical data
```{r write_table,eval=clinDB_need}
paste0("Patients with clinical data: ",length(unique(survival_data$britroc_number)))
paste0("Patients missing: ",paste0(which(!seq.int(1,276,1) %in% unique(survival_data$britroc_number)),collapse = ", "))
paste0("Duplicated patient rows: ",length(which(any(duplicated(survival_data$britroc_number)))))

## Final table
write.table(x = survival_data,file = "britroc_cohort_patient_data.tsv",quote = F,sep = "\t",row.names = F,col.names = T)
plot_data_db <- plot_data
```

## Summary plot
### Load surv data
```{r load_surv_data}
survival_data <- read.table(file = "britroc_cohort_patient_data.tsv",header = TRUE,sep = "\t",stringsAsFactors = F)
survival_data$britroc_number <- as.character(survival_data$britroc_number)

plot_data <- survival_data[order(survival_data$pt_sensitivity_at_reg,survival_data$os,decreasing = FALSE),]
diag_date <- plot_data
plot_data$britroc_number <- factor(plot_data$britroc_number,levels = plot_data$britroc_number)
plot_data$status <- ifelse(plot_data$status == 0, "alive","deceased")
```

### Get drugs/line data
```{r get_chemo_druglines,eval=clinDB_need}
plot_chemo_drugs <- RPostgres::dbReadTable(britroc_con, 'chemotherapy_lines_drugs')
plot_chemo_drugs$fk_britroc_number <- as.character(plot_chemo_drugs$fk_britroc_number)
plot_chemo_lines <- plot_chemo_drugs %>%
                    right_join(.,diag_date,by = c("fk_britroc_number"="britroc_number")) %>%
                    mutate(line_start=difftime(start_date,registration_date,units = "days")) %>%
                    mutate(line_end=difftime(completion_date,registration_date,units = "days")) %>%
                    mutate(britroc_number = fk_britroc_number)

plot_chemo_lines$britroc_number <- factor(plot_chemo_lines$britroc_number,levels = plot_data$britroc_number)
plot_chemo_lines <- plot_chemo_lines[order(plot_chemo_lines$britroc_number),]
plot_chemo_lines <- plot_chemo_lines[which(!is.na(plot_chemo_lines$chemotherapy_line)),]
plot_chemo_lines$chemotherapy_line[plot_chemo_lines$chemotherapy_line > 4] <- "5+"
```

```{r write_chemo_druglines,eval=clinDB_need}
write.table(plot_chemo_lines,file = "britroc_patient_treatment_data.tsv",quote = F,sep = "\t",row.names = F,col.names = T)
```

### Load drug data
```{r load_chem_data}
plot_chemo_lines <- read.table(file = "britroc_patient_treatment_data.tsv",
                               header = TRUE,
                               sep = "\t",stringsAsFactors = F)
plot_chemo_lines$britroc_number <- as.character(plot_chemo_lines$fk_britroc_number)
```

### Format plot data
```{r summary_plot_data}
plot_data_full <- plot_data %>%
      dplyr::select(britroc_number,pt_sensitivity_at_reg,status,os,int_start) %>%
      mutate(diag=0-int_start,os=os-int_start,int_start=int_start-int_start) %>%
      dplyr::select(-int_start) %>%
      pivot_longer(values_to= "time",names_to = "event",cols = c(diag,os))

plot_chemo_lines_full <- plot_chemo_lines %>%
                            dplyr::select(britroc_number,line_start,line_end,chemotherapy_line,pt_sensitivity_at_reg)
```

### Customise axis labels
```{r summary_plot_params}
year_breaks <- -seq(-730*2,730*12,730)
year_labs <- year_breaks / 365
year_labs[year_labs == 0] <- "Reg. date"
```

### Summary plot
```{r summary_plot}
sum_plot <- ggplot() +
              geom_hline(data = plot_data_full,aes(yintercept = britroc_number),color = "grey95") +
              geom_point(data = plot_data_full,aes(x = as.numeric(time),y = britroc_number,shape = status,color = event)) +
              scale_shape_manual(values = c(15,0)) +
              geom_point(data = plot_data_full[plot_data_full$event == "diag",],
                         aes(x = as.numeric(time),y = britroc_number,color = event),shape = 15) +
              scale_color_manual(values = c("deeppink2","#2171b5"),labels = c("diagnosis","last follow up")) +
              new_scale_color() +
              geom_linerange(data = plot_chemo_lines_full,
                             aes(y = britroc_number,xmin = as.numeric(line_start),xmax = as.numeric(line_end),color = chemotherapy_line),
                             size = 1,alpha = 0.8) +
              geom_vline(xintercept = 0) +
              scale_color_manual(values = colour_palettes$prior_lines) +
              ylab("Patient (grouped by platinum sensitivity)") +
              xlab("Years") +
              labs(title = "BriTROC cohort clinical timeline") +
              scale_x_continuous(expand = c(0,0),
                                 limits = c(min(plot_data_full$time)*1.05,max(plot_data_full$time)*1.05),
                                 breaks = year_breaks,labels = year_labs) +
              facet_grid(space = "free_y",scales = "free_y",rows = vars(pt_sensitivity_at_reg),switch = "y") +
              theme_bw() +
              theme(axis.text.y = element_blank(),
                    axis.ticks.y = element_blank(),
                    legend.position = "bottom",
                    panel.grid = element_blank())

sum_plot
saveRDS(object = sum_plot,file = "plots/cohort_plot.RDS")
ggsave(filename = "plots/cohort_summary_plot.png",plot = sum_plot,device = "png",width = 12,height = 9,dpi = 300,units = "in")
ggsave(filename = "plots/cohort_summary_plot.pdf",plot = sum_plot,device = "pdf",width = 12,height = 9,dpi = 300,units = "in")
```

### write source data
```{r source_data_export_1B}
write.table(x = plot_data_full,file = "source_data/figure_1B_layer_1.tsv",quote = F,sep = "\t",row.names = F,col.names = T)
write.table(x = plot_chemo_lines_full,file = "source_data/figure_1B_layer_2.tsv",quote = F,sep = "\t",row.names = F,col.names = T)
```

### Plot survival curve
```{r os_pt_status_surv}
## Survival from date of registration stratified by pt_status
survival_data_plot <- survival_data %>%
                    rename("platinum_status"="pt_sensitivity_at_reg") %>%
                    dplyr::select(britroc_number,int_start,int_end,status,platinum_status)
    
fit <- survfit(Surv(int_end - int_start, status) ~ platinum_status, data = survival_data_plot)
surv_plot <- ggsurvplot(fit,
           conf.int = TRUE,
           legend = "bottom",
           legend.title = "platinum status",
           xlab = "time (days)",
           ylab = "survival probability",
           legend.labs = names(colour_palettes$resist_sensitive),
           palette = as.character(colour_palettes$resist_sensitive),
           ggtheme = theme_bw())

surv_plot

saveRDS(object = surv_plot$plot,file = "plots/reg_surv_plot.RDS")
ggsave(filename = "plots/reg_surv_plot.png",plot = surv_plot$plot,device = "png",width = 8,height = 8,dpi = 300,units = "in")
ggsave(filename = "plots/reg_surv_plot.pdf",plot = surv_plot$plot,device = "pdf",width = 8,height = 8,dpi = 300,units = "in")
```

### Write source data
```{r source_data_export_1C}
write.table(x = survival_data_plot,file = "source_data/figure_1C.tsv",quote = F,sep = "\t",row.names = F,col.names = T)
```

## Session info
```{r sessioninfo}
sessionInfo()
```
