---
title: "Analysis of patient heterogeneity"
author: "Philip Smith"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

## Initial steps

- ANALYSIS of intra/inter group clonality (larger differences between spatial samples than temporal samples?)
- https://static-content.springer.com/esm/art%3A10.1038%2Fs41467-021-23384-6/MediaObjects/41467_2021_23384_MOESM1_ESM.pdf

### knitr options

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load libraries

```{r load_libraries}
# Load libraries
suppressPackageStartupMessages(suppressWarnings(library(Biobase)))
suppressPackageStartupMessages(suppressWarnings(library(QDNAseq)))
suppressPackageStartupMessages(suppressWarnings(library(QDNAseqmod)))
suppressPackageStartupMessages(suppressWarnings(library(tidyverse)))
suppressPackageStartupMessages(suppressWarnings(library(cowplot)))
suppressPackageStartupMessages(suppressWarnings(library(ggsignif)))
suppressPackageStartupMessages(suppressWarnings(library(survival)))
suppressPackageStartupMessages(suppressWarnings(library(survminer)))
suppressPackageStartupMessages(suppressWarnings(library(kableExtra)))
```

## Load data and funcs

Define functions specifically utilised in this document

### Functions
```{r funs}
## Get clonality / heterogeneity metric from absolute copy number profiles
get_clonality <- function(abs_data = NULL,weighted = TRUE,smooth=TRUE,wiggle=0.1,limit=1){
  # extract segments matrix
  segments <- assayDataElement(abs_data,elt = "segmented")
  
  # Weighted / smoothed method
  if(weighted){
    if(smooth){
      # Get length of segment vector as proxy for length of genome
      seg_len <- length(segments[,1])
      # set segments with values less than zero to zero
      segments[segments < 0] <- 0
      # Rounded segments matrix to whole integers
      rounded_segments <- round(segments)
      # Get boolean matrix of values which absolute difference
      # between unrounded and rounded segments is less than wiggle variable
      wiggleRemove <- which(abs(segments - rounded_segments) < wiggle)
      # Set TRUE boolean matrix values to NA in unrounded segment matrix
      segments[wiggleRemove] <- NA
      # between unrounded and rounded segments is greater than limit variable
      limitRemove <- which(abs(segments - rounded_segments) > limit)
      # Set TRUE boolean matrix values to NA in unrounded segment matrix
      segments[limitRemove] <- NA
      # Get segment rle object with NA values removed
      segment_rle <- apply(segments, 2, FUN = function(x) rle(x[which(!is.na(x))]))
      # Transform unrounded rle object values to absolute unrounded/rounded differences
      diff_rle <- lapply(segment_rle,FUN = function(x){
        x$values <- abs(x$values - round(x$values))
        return(x)
        })
      # Obtain weighted mean of absolute differences using number of bins per segments / total bins
      weightedMean <- lapply(diff_rle,FUN = function(x){
        wm <- weighted.mean(x = x$values,w = (x$lengths/seg_len),na.rm = TRUE)
        })
      # Named vector of weighted mean clonality value
      clonality <- unlist(weightedMean)
      return(clonality)
    } else if(!smooth){
      # Get length of segment vector as proxy for length of genome
      seg_len <- length(segments[,1])
      # set segments with values less than zero to zero
      segments[segments < 0] <- 0
      # Get segment rle object with NA values removed
      segment_rle <- apply(segments, 2, FUN = function(x) rle(x[which(!is.na(x))]))
      # Transform unrounded rle object values to absolute unrounded/rounded differences
      diff_rle <- lapply(segment_rle,FUN = function(x){
        x$values <- abs(x$values - round(x$values))
        return(x)
        })
      # Obtain weighted mean of absolute differences using number of bins per segments / total bins
      weightedMean <- lapply(diff_rle,FUN = function(x){
        wm <- weighted.mean(x = x$values,w = (x$lengths/seg_len),na.rm = TRUE)
        })
      # Named vector of weighted mean clonality value
      clonality <- unlist(weightedMean)
      return(clonality)
    }
  # Unweighted and unsmoothed method
  } else if(!weighted){
    if(smooth){
      # Get length of segment vector as proxy for length of genome
      seg_len <- length(segments[,1])
      # set segments with values less than zero to zero
      segments[segments < 0] <- 0
      # Rounded segments matrix to whole integers
      rounded_segments <- round(segments)
      # Get boolean matrix of values which absolute difference
      # between unrounded and rounded segments is less than wiggle variable
      wiggleRemove <- which(abs(segments - rounded_segments) < wiggle)
      # Set TRUE boolean matrix values to NA in unrounded segment matrix
      segments[wiggleRemove] <- NA
      # between unrounded and rounded segments is greater than limit variable
      limitRemove <- which(abs(segments - rounded_segments) > limit)
      # Set TRUE boolean matrix values to NA in unrounded segment matrix
      segments[limitRemove] <- NA
      # Get segment rle object with NA values removed
      segment_rle <- apply(segments, 2, FUN = function(x) rle(x[which(!is.na(x))]))
      # Transform unrounded rle object values to absolute unrounded/rounded differences
      diff_rle <- lapply(segment_rle,FUN = function(x){
        x$values <- abs(x$values - round(x$values))
        return(x)
        })
      # Obtain mean of absolute differences using number of bins per segments / total bins
      Mean <- lapply(diff_rle,FUN = function(x){
        m <- mean(x = x$values,na.rm = TRUE)
        })
      # Named vector of weighted mean clonality value
      clonality <- unlist(Mean)
      return(clonality)
    } else if(!smooth){
      # Get rounded segments
      int_segments <- round(segments)
      # Get col means of absolute differences between unrounded and rounded segments matrix
      clonality <- colMeans(abs(segments - int_segments),na.rm = TRUE)
      return(clonality)
    }
  }
}

get_cox_HR <-function(x){
  HR <- round(exp(coef(x)), 2)
  CI <- round(exp(confint(x)), 2)
  P <- round(coef(summary(x))[,5], 3)
  colnames(CI) <- c("Lower", "Higher")
  hazard_table <- as.data.frame(cbind(HR, CI, P))
  hazard_table <- cbind(Feature=row.names(hazard_table),hazard_table,row.names = NULL)
  return(hazard_table)
}

add_clonality_meta <- function(x){
  clonality_table <- data.frame(sample = names(x),clonality = x,row.names = NULL)
clonality_table$group <- meta.data$group[match(clonality_table$sample,meta.data$SAMPLE_ID)]
clonality_table$patient <- meta.data$PATIENT_ID[match(clonality_table$sample,meta.data$SAMPLE_ID)]
clonality_table$ploidy <- meta.data$ploidy[match(clonality_table$sample,meta.data$SAMPLE_ID)]
clonality_table$paired <- meta.data$paired[match(clonality_table$sample,meta.data$SAMPLE_ID)]
clonality_table$purity <- meta.data$purity[match(clonality_table$sample,meta.data$SAMPLE_ID)]
return(clonality_table)
}

add_corr <- function(x,type="limit"){
    x$zscore <- (x$clonality-mean(x$clonality,na.rm=T))/sd(x$clonality,na.rm = T)
    x$corrected <- x$zscore*x$ploidy*x$purity
    pl.clonality <- cor.test(x$ploidy,x$clonality,method = "kendal")$estimate
    pu.clonality <- cor.test(x$purity,x$clonality,method = "kendal")$estimate
    pl.zscore <- cor.test(x$ploidy,x$zscore,method = "kendal")$estimate
    pu.zscore <- cor.test(x$purity,x$zscore,method = "kendal")$estimate
    pl.corrected <- cor.test(x$ploidy,x$corrected,method = "kendal")$estimate
    pu.corrected <- cor.test(x$purity,x$corrected,method = "kendal")$estimate
    if(type == "limit"){
      limit_correlation <- rbind(limit_correlation,data.frame(limit=i,
                                   puclonality=pu.clonality,
                                   plclonality=pl.clonality,
                                   puzscore=pu.zscore,
                                   plzscore=pl.zscore,
                                   pucorrected=pu.corrected,
                                   plcorrected=pl.corrected,row.names = NULL))
      return(limit_correlation)
    } else if(type == "wiggle"){
      wiggle_correlation <- rbind(wiggle_correlation,data.frame(wiggle=i,
                                   puclonality=pu.clonality,
                                   plclonality=pl.clonality,
                                   puzscore=pu.zscore,
                                   plzscore=pl.zscore,
                                   pucorrected=pu.corrected,
                                   plcorrected=pl.corrected,row.names = NULL))
      return(wiggle_correlation)
    }
}

univar_surv <- function(surv = NULL,covariates = NULL,data = NULL){
  censor_surv <- surv 
  covariates <- covariates
  univ_formulas <- sapply(covariates,
                        function(x) as.formula(paste('censor_surv~', x)))
  univ_models <- lapply(univ_formulas, function(x){coxph(x, data = data)})
  univ_results <- lapply(univ_models,
                       function(x){ 
                          x <- summary(x)
                          p.value<-signif(x$wald["pvalue"], digits=2)
                          wald.test<-signif(x$wald["test"], digits=2)
                          beta<-signif(x$coef[1], digits=2);#coeficient beta
                          HR <-signif(x$coef[2], digits=2);#exp(beta)
                          HR.confint.lower <- signif(x$conf.int[,"lower .95"], 2)
                          HR.confint.upper <- signif(x$conf.int[,"upper .95"],2)
                          HR <- paste0(HR, " (", 
                                       HR.confint.lower, "-", HR.confint.upper, ")")
                          res<-c(beta, HR, wald.test, p.value)
                          names(res)<-c("beta", "HR (95% CI for HR)", "wald.test", 
                                        "p.value")
                          return(res)
                         })
  res <- t(as.data.frame(univ_results, check.names = FALSE))
  as.data.frame(res)
}

multivar_surv <- function(surv = NULL,covariates = NULL,data = NULL){
  censor_surv <- surv
  covariates <- covariates
  surv_formula <- as.formula(paste('censor_surv ~',paste(covariates,collapse = ' + ')))
  res.cox <- coxph(surv_formula,
                 data =  data)
  summary(res.cox)
  get_cox_HR(res.cox)
}

get_segment_noise <- function(data = NULL){
  if(is.null(data)){
    stop("no absolute cn data provided")
  }
  samples <- colnames(data)
  segment_noise <- do.call(rbind,lapply(samples,FUN = function(x){
    sample <- x
    s <- as.data.frame(assayDataElement(abs_data,"segmented"))
    b <- as.data.frame(assayDataElement(abs_data,"copynumber"))
    f <- fData(abs_data)[,1:3]
    
    s <- s[sample]
    b <- b[sample]
    
    t <- cbind(f,s,b)
    
    colnames(t) <- c("chromosome","start","end","segments","bin")
    tsd <- t %>%
      group_by(chromosome) %>%
      mutate(changepoint = abs(segments - lag(segments))) %>%
      mutate(change = ifelse(changepoint > 0.1 | is.na(changepoint),TRUE,FALSE)) %>%
      ungroup() %>%
      mutate(segment = cumsum(change)) %>%
      group_by(segment) %>%
      mutate(bin_count = n()) %>%
      mutate(bin_sd = sd(bin)) %>%
      mutate(seg_start = min(start)) %>%
      mutate(seg_end = max(end)) %>%
      mutate(segment_noise = weighted.mean(x = bin_sd,w = bin_count)) %>%
      ungroup() %>%
      mutate(sample = rep(sample,times=nrow(.)))
    return(tsd)
  }))
  return(segment_noise)
}
```

### Read meta data

Meta data for all samples with fitted absolute copy number profiles

```{r read_meta}
source("../../colour_palettes.R")
meta.data <- read.table("../../copy_number_signatures/britroc_30kb_signature_data_meta.tsv",
                        header = TRUE,
                        sep = "\t")

patient_data <- read.table("../../britroc_cohort_patient_data.tsv",
                           sep = "\t",
                           header = T)

patient_data$PATIENT_ID <- paste0("BRITROC-",patient_data$britroc_number)
```

### Read absolute data

Absolute copy number profiles are loaded in QDNAseq object format and unused/dropped samples are removed. 

```{r read_abs}
abs_data <- readRDS("../../absolute_POST_down_sampling/britroc_30kb_ds_absCopyNumber.rds")
abs_data <- abs_data[,which(colnames(abs_data) %in% meta.data$SAMPLE_ID[meta.data$use == "TRUE"])]
```

## Analysis

Amendment - No smoothing is now applied to either the upper or lower bounds applied to each segment to more closely conform with the approach seen by van Dijk et al.see [here](https://www.nature.com/articles/s41467-021-23384-6).

```{r segment_noise}
segment_noise <- get_segment_noise(data = abs_data)

## Values exceeding 2sd from mean excluded
segment_noise_segs <- segment_noise[segment_noise$change == TRUE,]

segment_threshold <- (mean(segment_noise_segs$segment_noise) +
                             sd(segment_noise_segs$segment_noise)*2)

segment_noise_segs <- segment_noise_segs %>%
                          mutate(excluded = ifelse(segment_noise >= segment_threshold,
                                                TRUE,FALSE))

segment_noise_sample <- segment_noise_segs %>%
                          group_by(sample) %>%
                          summarise(across(segment_noise,mean))

sample_threshold <- (mean(segment_noise_sample$segment_noise) +
                             sd(segment_noise_sample$segment_noise)*2)

segment_noise_sample <- segment_noise_sample %>%
                          mutate(excluded = ifelse(segment_noise >= sample_threshold,
                                                TRUE,FALSE))
```

```{r noise_plots}
ggplot(segment_noise_sample) +
  geom_histogram(aes(segment_noise,fill=excluded),binwidth = 0.04) +
  geom_vline(xintercept = sample_threshold,
             linetype=2) +
  xlab("sample noise") +
  ylab("") +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  theme_bw()

ggplot(segment_noise_segs) +
  geom_histogram(aes(segment_noise,fill=excluded),binwidth = 0.1) +
  geom_vline(xintercept = segment_threshold,
             linetype=2) +
  xlab("segment noise") +
  ylab("") +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  theme_bw()

ggplot(segment_noise_segs) +
  geom_point(aes(segments,segment_noise,colour=excluded)) +
  xlab("copy number") +
  ylab("segment noise") +
  theme_bw()

ggplot(segment_noise_segs) +
  geom_bar(aes(excluded,fill=excluded)) +
  xlab("excluded") +
  ylab("segments") +
  theme_bw()

```

```{r noise_stats}
table(segment_noise_segs$excluded) / nrow(segment_noise_segs) * 100
segment_stats <- segment_noise_segs %>%
                  group_by(sample) %>%
                  mutate(seg_n = max(segment)) %>%
                  mutate(exc_segs = sum(excluded == TRUE)) %>%
                  mutate(exc_pct = exc_segs/seg_n*100) %>%
                  dplyr::select(sample,seg_n,exc_segs,exc_pct) %>%
                  distinct()
summary(segment_stats$exc_segs)
```

### Clonality analysis
```{r apply_sample_filter}
excluded_samples <- segment_noise_sample$sample[segment_noise_sample$excluded == TRUE]
excluded_paired_sample <- as.character(meta.data$SAMPLE_ID[meta.data$SAMPLE_ID %in% excluded_samples & meta.data$paired == TRUE])

cat("Excluded samples: ",paste0(excluded_samples,collapse = ",")," (n=",length(excluded_samples),")",sep = "")
abs_data_filt <- abs_data[,which(!colnames(abs_data) %in% excluded_samples)]
```
```{r noise_stats_post_filt}
segment_noise_segs_filt <- segment_noise_segs[which(!segment_noise_segs$sample %in% excluded_samples),]
table(segment_noise_segs_filt$excluded) / nrow(segment_noise_segs_filt) * 100
segment_stats_filt <- segment_noise_segs_filt %>%
                  group_by(sample) %>%
                  mutate(seg_n = max(segment)) %>%
                  mutate(exc_segs = sum(excluded == TRUE)) %>%
                  mutate(exc_pct = exc_segs/seg_n*100) %>%
                  dplyr::select(sample,seg_n,exc_segs,exc_pct) %>%
                  distinct()
summary(segment_stats_filt$exc_segs)
```

```{r apply_seg_filter}
segment_noise_filt <- as.data.frame(segment_noise %>%
                  mutate(exclude_seg = ifelse(segment_noise >= segment_threshold,TRUE,FALSE)) %>%
                  dplyr::select(sample,chromosome,start,end,segments,exclude_seg))
                  
segment_noise_filt <- split(segment_noise_filt,f = segment_noise_filt$sample)
segment_noise_filt <- segment_noise_filt[which(!names(segment_noise_filt) %in% excluded_samples)]

if(!identical(names(segment_noise_filt),colnames(abs_data_filt))){
  stop("sample names do not match")
}

samples <- names(segment_noise_filt)
for(x in samples){
  #print(x)
  segment_filt_sample <- segment_noise_filt[[x]]
  
  if(!all(segment_filt_sample$segments == assayDataElement(object = abs_data_filt,"segmented")[,x])){
    stop(paste0(x," had non-matching segment values - seek help."))
  }
  if(sum(segment_filt_sample$exclude_seg) > 0){
    assayDataElement(object = abs_data_filt,"segmented")[which(segment_filt_sample$exclude_seg == TRUE),x] <- NA
  }
}
```

```{r set_clonality}
clonality_table <- add_clonality_meta(get_clonality(abs_data = abs_data_filt,
                             weighted = TRUE,
                             smooth = FALSE,
                             wiggle = 0.1)) %>%
                    mutate(group = case_when(group == "arx" ~ "diagnosis",group == "rlps" ~ "relapse")) %>%
                    rename(ith="clonality")

kableExtra::kable(clonality_table) %>%
  kable_styling() %>%
  scroll_box(height = "200px")
write.table(clonality_table,file = "data/ith_results_table.tsv",quote = F,row.names = F,sep = "\t",col.names = T)
```

```{r plot_clonaltiy_basics}
ggplot(clonality_table) +
  geom_density(aes(ith)) +
  theme_bw()

ggplot(clonality_table) +
  geom_point(aes(ploidy,ith)) +
  geom_smooth(aes(ploidy,ith),method = "lm") +
  theme_bw()

cor.test(clonality_table$ploidy,clonality_table$ith,method = "kendal")

ggplot(clonality_table) +
  geom_point(aes(purity,ith)) +
  geom_smooth(aes(purity,ith),method = "lm") +
  theme_bw()

cor.test(clonality_table$purity,clonality_table$ith,method = "kendal")
```

```{r clonality_by_type}
ggplot(clonality_table) +
  geom_point(aes(group,ith,fill=group),position="jitter",color="grey50") +
  geom_violin(aes(group,ith,fill=group),alpha=0.5) +
  geom_signif(aes(group,ith),comparisons = list(c("diagnosis","relapse")),test = "wilcox.test") +
  scale_fill_manual(values = colour_palettes$diagnosis_relapse) +
  theme_bw()

excluded_pats <- as.character(meta.data$PATIENT_ID[meta.data$SAMPLE_ID %in% excluded_samples & meta.data$paired == TRUE])

clonality_table_paired <- clonality_table %>%
                            filter(!sample %in% meta.data$SAMPLE_ID[meta.data$PATIENT_ID %in% excluded_pats]) %>%
                            filter(paired == TRUE) %>%
                            group_by(patient,group) %>%
                            summarise(across(.cols = c(ploidy,purity,ith),.fns = median))
                            

clonality_plot_paired <- ggplot(clonality_table_paired) +
  geom_point(aes(group,ith,fill=group),color="grey50") +
  geom_line(aes(group,ith,group=patient),alpha=0.5) +
  geom_violin(aes(group,ith,fill=group),alpha=0.5) +
  geom_signif(aes(group,ith),comparisons = list(c("diagnosis","relapse")),test = "wilcox.test",test.args = list(paired=TRUE)) +
  scale_fill_manual(values = colour_palettes$diagnosis_relapse) +
  theme_bw()

ggsave(filename = "plots/paired_clonality_plot.png",plot = clonality_plot_paired,width = 9,height = 7,units = "in",dpi = 300)
clonality_plot_paired
```

#### Intra-sampe clonality

Clonality values were not different at a cohort level between primary and relapse tumours but did have variation within the groups.

```{r clonality_arx_data}
arx_clonality_table <- clonality_table %>%
                        filter(group == "diagnosis") %>%
                        left_join(.,patient_data,by = c("patient"="PATIENT_ID"))
```

### Clonality change

```{r patient_changes}
patient_clonality_diffs <- clonality_table %>%
                            filter(!sample %in% meta.data$SAMPLE_ID[meta.data$PATIENT_ID %in% excluded_pats]) %>%
                            dplyr::filter(paired == "TRUE") %>%
                            group_by(patient,group) %>%
                            summarise(ith=mean(ith)) %>%
                            ungroup() %>%
                            group_by(patient) %>%
                            mutate(ith_change = ith[group == "relapse"] - ith[group == "diagnosis"]) %>%
                            dplyr::select(patient,ith_change) %>%
                            distinct()
```

```{r clonality_change_dist}
ggplot(patient_clonality_diffs) +
  geom_density(aes(ith_change)) +
  theme_bw()
```

```{r patient_ploidy}
patient_ploidy <- meta.data %>%
  group_by(PATIENT_ID) %>%
  summarise(ploidy=mean(ploidy))

patient_purity <- meta.data %>%
  group_by(PATIENT_ID) %>%
  summarise(purity=mean(purity))
```

```{r clonality_diff_cors}
patient_clonality_diffs <- patient_clonality_diffs %>%
  inner_join(.,patient_data,by = c("patient"="PATIENT_ID")) %>%
  mutate(.data = .,diagnosis_distance = as.numeric(difftime(diagnosis_date,as.Date(format(Sys.time(), '%Y-%m-%d')))))

patient_clonality_diffs$ploidy <- patient_ploidy$ploidy[match(patient_clonality_diffs$patient,patient_ploidy$PATIENT_ID)]
patient_clonality_diffs$purity <- patient_purity$purity[match(patient_clonality_diffs$patient,patient_purity$PATIENT_ID)]
```

```{r clonality_diff_age}
ggplot(patient_clonality_diffs) +
  geom_point(aes(age,ith_change)) +
  geom_smooth(aes(age,ith_change),method = "lm") +
  theme_bw()

cor.test(patient_clonality_diffs$age,patient_clonality_diffs$ith_change,method = "kendal")
```

```{r clonality_diff_sample_age}
ggplot(patient_clonality_diffs) +
  geom_point(aes(diagnosis_distance,ith_change)) +
  geom_smooth(aes(diagnosis_distance,ith_change),method = "lm") +
  theme_bw()

cor.test(patient_clonality_diffs$diagnosis_distance,patient_clonality_diffs$ith_change,method = "kendal")
```

```{r clonality_diff_sensitivity}
delta_itg_pt_sens <- ggplot(patient_clonality_diffs) +
  geom_point(aes(pt_sensitivity_at_reg,ith_change,fill=pt_sensitivity_at_reg),position="jitter") +
  geom_violin(aes(pt_sensitivity_at_reg,ith_change,fill=pt_sensitivity_at_reg),alpha=0.6) +
  geom_signif(aes(pt_sensitivity_at_reg,ith_change),comparisons = list(c("resistant","sensitive")),test = "wilcox.test") +
  ylab(label = paste0("\u0394","ITH")) +
  xlab("") +
  scale_fill_discrete(name = "platinum status") +
  theme_bw()
delta_itg_pt_sens
ggsave(filename = "plots/delta_itg_pt_sens.png",plot = delta_itg_pt_sens,width = 8,height = 6,units = "in",dpi = 300)
```

```{r clonality_diff_priorlines}

# delta_itg_priorlines <- 
patient_clonality_diffs$pre_reg_chemo <- as.factor(patient_clonality_diffs$pre_reg_chemo)
  
  ggplot(patient_clonality_diffs) +
  geom_point(aes(pre_reg_chemo,ith_change,fill=pre_reg_chemo),position="jitter") +
  geom_violin(aes(pre_reg_chemo,ith_change,fill=pre_reg_chemo),alpha=0.6) +
  geom_signif(aes(as.factor(pre_reg_chemo),ith_change),comparisons = list(c("1","2")),test = "wilcox.test") +
  ylab(label = paste0("\u0394","ITH")) +
  xlab("") +
  scale_fill_discrete(name = "prior lines") +
  theme_bw()
delta_itg_pt_sens
ggsave(filename = "plots/delta_itg_pt_sens.png",plot = delta_itg_pt_sens,width = 8,height = 6,units = "in",dpi = 300)
```

##### Clonality patient-level analysis

```{r arx_multi_sample_data}
multisite_patients <- clonality_table %>%
  group_by(patient) %>%
  tally() %>%
  filter(n > 1) %>%
  dplyr::select(patient)

multisite_data <- clonality_table %>%
  filter(patient %in% multisite_patients$patient)

plot(as.dendrogram(
  hclust(
    dist(x = as.data.frame(multisite_data[multisite_data$patient == "BRITROC-232",c("ith")],
                           row.names = as.character(multisite_data[multisite_data$patient == "BRITROC-232",c("sample")]))))))
                
```


